---
title: "What are the results of our methods for defining events?"
format: html
editor: visual
---

## Setup

```{r}
#| output: false

library(identidrift)
library(tidyverse)
```

## Intro

In the methods we outlined the process for how we defined events.

1.  The first part of that was to use a template click with the MTC module to generate match scores for all clicks in each database.
2.  After that we used a custom script to construct the events. The script output some data that is consolidated into the `template` data object.
3.  All the events that were defined in step 2 were then used to train the classifier. This final selection of data used to train the model is available in the `nbhf_clicks` data object

## Questions

-   Did our methods fail to define any events, that should have been included i.e. did it result in any **False Negatives**?
    -   Approach: evaluate distribution of match scores for all clicks that are **not** included in events that made it into the training set.
    -   Use `template` data to plot this distribution. See @fig-template.
-   Did our methods identify any **False Positives**?
    -   This is unlikely as the thresholds decided upon were chosen to set an intentionally low bar, and we performed a manual review to determine that no manually labelled events were being excluded from the automatically defined events when choosing these thresholds.

The end result of that process

Some questions that we could answer to give insight into our results

```{r}
glimpse(template)
```

You can see, this output include for each UID:

-   match scores, across all templates (`_thresh` columns appear to be copies of `_match` columns)
-   a `type` field which I think is related to the click classifier label given in PG

What can we glean from this `template` data to give insight into our results?

### Comparing distribution of match scores in and outside of events

**Clicks not included in events are, by and large, poor template matches.**

I would like to make some comparisons in the distribution of match scores across clicks that were included in events, versus those that did not get included.

First thing we need to do to is to determine how we distinguish between clicks that are inside or outside of an event, in the `template` dataset.

Since all the events are included in the training set, we can compare UIDs in the training set object called `nbhf_clicks` to the UIDs in `template` to determine identifying marks.

The number of clicks in the training data set is `{r} nrow(nbhf_clicks)` and the number of clicks in the `template` dataset having `type==101` is `{r} nrow(template%>%filter(type==101))`. There is obviously a small discrepancy here but, overall the numbers match up pretty well. Consider the quantity of clicks broken down by drift.

::: panel-tabset
## template type=101

```{r}
#| echo: false

template %>%
  filter(type==101) %>%
  count(db)
```

## training

```{r}
#| echo: false
#| 
nbhf_clicks %>%
  count(db) %>%
  mutate(db=str_remove(basename(db), " - Copy.sqlite3"))
```
:::

This leads me to conclude that filtering the `template` data for clicks with `type==101` is shorthand for filtering clicks included in events.

### Distributions

```{r}

# First thing we will do to evaluate distributions is subset `template` so that we can distinguish between clicks inside events and those not inside events

t <- template %>% 
  mutate(inev=case_when(type==101 ~ TRUE, .default=FALSE))

# Then get the best match score for each click across all the templates

t <- t %>%
  select(db, UID, type, ends_with("thresh"), inev) %>% 
  mutate(best=pmax(template$Pd_1_thresh, template$Pd_2_thresh, template$Pp_thresh,
     template$Ksp_thresh))
  
# Then join with species and eventId so we can look at distributions across individual events

j1 <- nbhf_clicks %>%
  distinct(species, db) %>%
  mutate(db = str_remove(basename(db), " - Copy.sqlite3"))

t <- t %>%
  left_join(j1, by="db")

t <- nbhf_clicks %>%
  select(UID, eventId) %>%
  right_join(t, by="UID")
```

These distributions show, probably with a significant difference, that clicks not included in events have lower match scores than those included in events. Suggesting that the script did a good job of placing highly matched clicks in events.

```{r}
#| echo: false
  t %>% 
    ggplot(aes(x=inev, y=best))+
    geom_violin()
```

There are however some high-scoring clicks that did not get included into events, likely because they were isolated and thus did not meet the criteria for an event. To get a sense of just how many high quality clicks were excluded, lets look at the percentile for the threshold value among the not-included clicks

```{r}
#| echo: false
vals <- data.frame(species = c("pd", "pp","ks"),
                   threshVal = c(0.45,0.6,0.45))



#use threshold value for each species in val table above to determine the percentile of the value among all out-of-event clicks

excl <- t %>%
  filter(inev==FALSE) %>% 
  nest(data=-species) %>% 
  inner_join(vals, by = "species") %>%
  mutate(pctile = map_dbl(.x=data,
                          .f= ~100-length(.x$best[.x$best < threshVal])/length(.x$best)*100))
  
```

We see from this that

-   for harbor porpoise, `{r} excl[excl$species=="pp","pctile"]`% of non-event clicks were above the threshold.
-   for Dall's porpoise, `{r} excl[excl$species=="pd","pctile"]`% of non-event clicks were above the threshold.
-   for *Kogia*, `{r} excl[excl$species=="ks","pctile"]`% of non-event clicks were above the threshold.
