---
title: "Thesis Results Notebook"
author: Jackson Vanfleet-Brown
date: 02/08/2024
format: html
editor: source
echo: false
toc: true
toc-depth: 3
---

::: callout-note
## Notebook organization

Results are organized in association with the methodological step which generated them.

The Table of Contents reflects this organizational scheme
:::

```{r}
#| output: false

library(identidrift)
library(tidyverse)
library(PAMpal)
library(densityClust)
library(rfPermute)
library(seewave)
```

# 1. Making Events

First step of the study was to define events. This is associated with three constituent methodologies.

1.  Selection of template clicks @sec-templates
2.  Using PG to generate MTC scores @sec-MTC
3.  Using custom R script to define events @sec-customR

## Template selection {#sec-templates}

### Methodology

-   Chose template clicks representing high-quality clicks
-   All are saved in the `identidrift::click_templates` data object @sec-data-template
-   Chose 4 total, 2 for Dall's porpoise and 1 for each other species. Two were needed for Dall's porpoise due to apparent increased variability in events.

### Results

I can add some general remarks here about broad differences in the shapes of the different profiles and what features led us to select them as high-quality clicks.

```{r}
#| label: fig-click-template-timewaves
#| fig-cap: "Template click time waves."
#| fig-subcap: 
#|   - "Kogia"
#|   - "Dall's porpoise 1"
#|   - "Dall's porpoise 2"
#|   - "Harbor porpoise"
#| layout-ncol: 2

plot(click_templates[["Ksp"]]$wave[,2], type = 'l', xlab = "Time", ylab = "Amplitude")
plot(click_templates[["Pd_1"]]$wave[,1], type = 'l', xlab = "Time", ylab = "Amplitude")
plot(click_templates[["Pd_2"]]$wave[,1], type = 'l', xlab = "Time", ylab = "Amplitude")
plot(click_templates[["Pp"]]$wave[,2], type = 'l', xlab = "Time", ylab = "Amplitude")

```

```{r}
#| label: fig-click-template-spectra
#| fig-cap: "Template click spectra"
#| fig-subcap: 
#|   - "Kogia"
#|   - "Dall's porpoise 1"
#|   - "Dall's porpoise 2"
#|   - "Harbor porpoise"
#| layout-ncol: 2

# Could also use PAMpal::plotSpectrogram() !!!
spec(click_templates[["Ksp"]]$wave, 384000)
spec(click_templates[["Pd_1"]]$wave, 384000)
spec(click_templates[["Pd_2"]]$wave, 384000)
spec(click_templates[["Pp"]]$wave, 384000)

```
### Discussion points

This is important for those who are interested in seeing the exact profiles of these types of signals and could be helpful for people studying NBHF clicks in other regions, to compare to what we consider to be a high-quality click.

## Template matching {#sec-MTC}

### Methodology

-   Used *PAMguard* MTC module to generate match scores for all database clicks.
-   The **best** match score along with the associated template ID is saved for each click in the `identidrift::template` data object.

### Results

From @fig-best-match-score-dist we can look comparatively at how each species' clicks matched against the four available templates.

The distributions show that:

 - *Kogia* seems to match best with it's template, since the distribution mean is well above the other two. This indicates also that the other species don't match well with it.
 - For the harbor porpoise template, the flattened distribution of harbor porpoise appears to have a higher mean than the other species. Another indication that this template is working well, like the *Kogia* template
 - Mixed results for Dall's porpoise. Not seeing the same trends that would indicate that it is being matched well with it's templates.

```{r}
#| label: fig-best-match-score-dist
#| fig-cap: Species distributions of best match scores, faceted by template

template %>%
  ggplot(aes(x=value, color=species))+
  geom_density()+
  facet_wrap(~best)
```

### Discussion points

- Harbor porpoise and *Kogia* templates is doing a good job of matching with their respective species while excluding others.
- Will the poorer matching of Dall's porpoise be reflected in the classifier's results as well? This could say something about this species, such as that it exhibits greater variability and has a signature that is harder to constrain.

## Define events {#sec-customR}

### Methodology

-   Used custom script to label the events in each database. Required a match score threshold to be given for each template (0.45-0.6)
- Also required setting maximum separation between clicks in the same event, given as **120 sec**, and minimum number of clicks per event, given as **3**.
- Script then labeled events in the database
    - Only clicks that were seen as part of an event were included in the training data set.
    - Clicks not included in events were not used for any further analysis.

### Results

Recall and precision would be common metrics for assessing the performance of our event definition procedure. 
Recall (@tbl-recall) would be estimated by counting clicks that possessed high match scores but which did not make it into the training set (false negatives). This would occur when high-quality clicks were isolated and did not occur temporally with enough other clicks to warrant declaring an event. This means they will have been ommitted from the training set. This does appear to be a rather small component since as we see in @tbl-recall recall is around ~90% across all species.

```{r}

dropped <- setdiff(template$UID, nbhf_clicks$UID)
recall <- template[which(template$UID %in% dropped),] %>% 
  filter(value > 0.55) %>% # crude filter, need to improve, use nest/mutate/map
  group_by(species) %>%
  count() %>% 
  inner_join(count(nbhf_clicks, species), by="species") %>%
  mutate(recall=n.y/(n.y+n.x)*100)

```

|        Ksp          |          Pd         |        Pp           |
|---------------------|---------------------|---------------------|
| `{r} recall[1, 4]`% | `{r} recall[2, 4]`% | `{r} recall[3, 4]`% |

: Recall of event definition procedure {#tbl-recall}

Precision would be more difficult to estimate. There is a region of overlap in the score distributions of the data that was and was not included (see @fig-match-score-overlap). This suggests sort of a "grey" area where the match is sometimes poor, but with enough other matches could nonetheless warrant an event?
```{r}
#| label: fig-match-score-overlap
#| fig-cap: Distributions of scores of clicks both included and excluded from the training set.

template %>%
  mutate(incl = case_when(UID %in% nbhf_clicks$UID ~ TRUE,
                          .default = FALSE)) %>%
  ggplot()+
  geom_density(aes(x=value, fill=incl), alpha=0.4) +
  facet_wrap(~species)

```
To investigate further I plot to see if there is a relationship between the mean match score of an event and the size of an event. @fig-score-vs-size shows that larger events tend to have higher mean match scores. Looks like a pretty strong positive relationship.
 
```{r}
#| label: fig-score-vs-size

template %>%
  inner_join(nbhf_clicks, by="UID") %>%
  group_by(eventId) %>%
  summarize(n=n(),
            mean=mean(value)) %>%
  ggplot()+
  geom_point(aes(x=n, y=mean))+
  scale_x_log10()
```

### Discussion points

- Recall is high, which we expected because we deliberately lowered the threshold out of concern for missing valuable data in our training set.
- Precision is harder to estimate. We might expect precision to have been diminished since we tailored our model to maximize recall.

# 2. Clustering

Next step of the study was to cluster to discover associations between events and/or clicks indicative of species classification. Broken down into two steps:

1.  Use PAMpal to extract click features @sec-PAMpal
2.  Cluster based on event @sec-ev-clusters
3.  Density cluster based on clicks @sec-density-clusters
4.  Density cluster, within species, based on clicks @sec-density-clusters-sp

## Extracting features {#sec-PAMpal}

### Methodology

-   From the database with the training events labeled, we then used PAMpal to extract standard set of features from each click.
-   What are the average or generalized spectra and concatenated spectrogram for each different species? In other words, what are the patterns that unify each species and distinguish it from other species?
-   What are the distributions of the standard click features across all the groups?
-   The training data is saved in the `nbhf_clicks` object @sec-data-nbhf_clicks

### Results

```{r}
#| label: fig-ks-concat
#| fig-cap: "Spectral features of Kogia species class"
#| fig-subcap: 
#|   - "Concatenated spectrogram"
#|   - "Averaged spectrum"
#| layout-ncol: 2

calculateAverageSpectra(nbhf_studies[["ks"]], evNum=1:length(events(nbhf_studies[["ks"]])))
```

```{r}
#| label: fig-pd-concat
#| fig-cap: "Spectral features of Dall's porpoise class"
#| fig-subcap: 
#|   - "Concatenated spectrogram"
#|   - "Averaged spectrum"
#| layout-ncol: 2

calculateAverageSpectra(nbhf_studies[["pd"]], evNum=1:length(events(nbhf_studies[["pd"]])))
```

```{r}
#| label: fig-pp-concat
#| fig-cap: "Spectral features of harbor porpoise class"
#| fig-subcap: 
#|   - "Concatenated spectrogram"
#|   - "Averaged spectrum"
#| layout-ncol: 2

calculateAverageSpectra(nbhf_studies[["pp"]], evNum=1:length(events(nbhf_studies[["pp"]])))
```

## Cluster based on event {#sec-ev-clusters}

### Methodology

-   Used nmds with community ecology tactics

### Results

-   Does clustering suggest that there are differences between species?
-   Does clustering suggest that, for an individual species, there are differences between events?

## Density Cluster clicks {#sec-density-clusters}

### Methodology

-   Used `densityClust` with `rho` and `delta` chosen in advance or manually chosen?

### Results

-   Does clustering suggest that there are differences between species?

## Density Cluster clicks, by species {#sec-density-clusters-sp}

### Methodology

-   Used `densityClust` with `rho` and `delta` chosen in advance or manually chosen?

### Results

-   Does clustering suggest that, for an individual species, there are differences between events?

# Model classifier

**IN PROGRESS**

# Data Appendix

## `nbhf_clicks` {#sec-data-nbhf_clicks}

## `template` {#sec-data-template}

You can see, this data includes for each UID (`{r} nrow(template)` total):

-   match scores, across all 4 templates
-   `_thresh` columns appear to be copies of `_match` columns
-   a `type` field which I think is related to the click classifier label given in PG
-   The number of clicks in the training data set is `{r} nrow(nbhf_clicks)` and the number of clicks in the `template` dataset having `type==101` is `{r} nrow(template%>%filter(type==101))`. There is obviously a small discrepancy here but still leads me to conclude that filtering the `template` data for clicks with `type==101` is shorthand for filtering clicks included in events.
-   This can be verified by examining counts, by event

```{r}
#| label: tbl-template
glimpse(template)
```
