---
title: "NBHF event clusters analysis"
author: Jackson Vanfleet-Brown
bibliography: references.bib
---

# Clustering Acoustic Events

```{r setup}
#| output: false

library(vegan)
library(tidyverse)
library(identidrift)
library(PAMpal)

s <- bindStudies(pd_mtc, pp_mtc, ks_mtc)
c <- getClickData(s) %>% best_clicks() #get tidy data
```

Our objective in this study is to classify acoustic events to species. Events we define as bouts of clicks that occur when an animal or group of animals encounters the recorder. We hypothesize that click features depend on the species that produced them, so what advantage do we give ourselves by seeking to classify events rather than individual clicks? The advantage of classifying events is the ability to overcome poor classification performance at the click level--see Rankin et al. [-@rankin2017] for an explanation. Briefly, it's because the distribution of click classifications within an event is indicative of the correct classification of the event as a whole.

Before trying to perform event classification, we decided we would perform event clustering to assess separation between events and species. One way of clustering these events would be to treat them as communities.

## Acoustic events as "communities" of clicks

In ecological research, communities are compared to one another by counting the species that appear in each community. This enables the generation of a dissimilarity index between different communities. When applied to plant communities, its easy to picture how this sort of counting would be performed. It's less clear in the case of a microbial community, for example, where we rely on genome sequence data to draw comparisons. In this case, a variety of different operational taxonomic units are defined and the quantity of sequences matching each operational unit are counted to characterize the community. This is analogous to the approach that can be taken with acoustic data when comparing acoustic events.

What needs to be accomplished is essentially a binned counting of the clicks in each event. This binning is accomplished by subdividing the region of the acoustic feature space in which we expect there to be divisions among the different species. For example, if we wanted to compare events on the basis of peak frequency, we could subdivide in bins the region of the acoustic spectrum from 100-160 kHz. If we chose breaks every 1 kHz, then seperate bins would exist for clicks with a peak frequency 100-101 kHz, 101-102 kHz, ... 159-160 kHz. The dissimilarity index between two events, in terms of their peak frequency, would then be generated by counting the number of clicks in each bin.

@fig-bins illustrates how binning and counting clicks would be performed across all events for each of the three focal species. The height of the bar in each bin represents the quantity of clicks produced by that species, in the available recordings, with a peak frequency in the operational unit as defined on the x-axis. You can imagine how it would be possible to generate a dissimilarity index among the three species using these counts.

```{r}
#| label: fig-bins
c %>%
  ggplot(aes(x=peak))+
  geom_histogram(binwidth=1)+
  facet_wrap(~species, ncol=1)
```
The current objective however, is to cluster events and then determine if species differences cause separation in events.

```{r bin}
pk <- eventbin(c, peak) #perform binning procedure
bw <- eventbin(c, BW_3dB)
ctr <- eventbin(c, centerkHz_3dB)
```

```{r clusters}
#| message: false
clust <- function(mat, title){
  dist <- vegdist(mat, method = "bray")
  nmds <- metaMDS(dist)
  scores(nmds) %>%
    as_tibble(rownames = "eventId") %>%
    left_join(select(c, species, eventId, eventLabel), by="eventId") %>% 
    ggplot(aes(x = NMDS1, y = NMDS2, color=species))+
    geom_point()+
    # facet_wrap(~eventLabel) +
    ggtitle(title)
}

set.seed(123)
clust(pk, "peak")
clust(bw, "bandwidth 3dB")
clust(ctr, "centerfreq 3dB")
```

Other things to consider: a null model? Rarefaction or sampling so that all events are the same size before clustering?

```{r null}
null <- mutate(c, eventId=sample(eventId))
npk <- eventbin(null, peak) #perform binning procedure
nbw <- eventbin(null, BW_3dB)
nctr <- eventbin(null, centerkHz_3dB)
clust(npk, "null peak")
clust(nbw, "null BW")
clust(nctr, "null center freq")
```
