---
title: "xtra"
format: html
editor: source
---

## Data available

### Match scores

#### Clicks in events used for training

```{r}
glimpse(nbhf_clicks)
```

## Clicks in/not in events

First thing we need to do to is to determine how we distinguish between clicks that are inside or outside of an event, in the `template` dataset.

Since all the events are included in the training set, we can compare UIDs in the training set object called `nbhf_clicks` to the UIDs in `template` to determine identifying marks.

The number of clicks in the training data set is `{r} nrow(nbhf_clicks)` and the number of clicks in the `template` dataset having `type==101` is `{r} nrow(template%>%filter(type==101))`. There is obviously a small discrepancy here but, overall the numbers match up pretty well. Consider the quantity of clicks broken down by drift.

::: panel-tabset
## template type=101

```{r}
#| echo: false

template %>%
  filter(type==101) %>%
  count(db)
```

## training

```{r}
#| echo: false
#| 
nbhf_clicks %>%
  count(db) %>%
  mutate(db=str_remove(basename(db), " - Copy.sqlite3"))
```
:::

### Matches

In @fig-best-match-score-bars we can look at just clicks exceeding the chosen threshold. We see what proportion of each species' true positive clicks (those having a score exceeding threshold) matched with each of the four templates. We see that the overwhelming majority of harbor porpoise clicks matched with the intended template. While both *Kogia* and Dall's porpoise appear to have about half of their clicks matched with the intended template, they also have a substantial proportion of clicks matching well with other templates.

```{r}
#| label: fig-best-match-score-bars
#| echo: false
#| fig-cap: Proportion of clicks with match scores over threshold belonging to each species, colored by template

template %>%
  filter(value>0.5) %>% 
  ggplot(aes(x=species, fill=best))+
  geom_bar(position="fill")
```


```{r}
#| echo:false

x <- template %>%
  group_by(species) %>% 
  count

y <- nbhf_clicks %>%
  count(species)

z <- inner_join(x, y, by="species") %>%
  mutate(pct_reduced=100-(n.y/n.x)*100)
```

The thresholding process and event definition process, with the values chosen, reduced the size of the databases appreciably with the harbor porpoise data being reduced by `{r} z[3, 4]`%, *Kogia* database reduced by `{r} z[1, 4]`%, and Dall's porpoise reduced by `{r} z[2, 4]`%.

```{r}
#| eval: false

nbhf_clicks %>%
  count(species, eventId) %>%
  arrange(n) %>% 
  ggplot(aes(x=eventId, y=n, color=species))+
  geom_bar(stat="identity")
```

Thresholds chosen reduced size of databases by how much percent? i.e. how selective were we being?
