---
title: "xtra"
format: html
editor: visual
---

## Data available

### Match scores

#### Clicks in events used for training

```{r}
glimpse(nbhf_clicks)
```

## Clicks in/not in events

First thing we need to do to is to determine how we distinguish between clicks that are inside or outside of an event, in the `template` dataset.

Since all the events are included in the training set, we can compare UIDs in the training set object called `nbhf_clicks` to the UIDs in `template` to determine identifying marks.

The number of clicks in the training data set is `{r} nrow(nbhf_clicks)` and the number of clicks in the `template` dataset having `type==101` is `{r} nrow(template%>%filter(type==101))`. There is obviously a small discrepancy here but, overall the numbers match up pretty well. Consider the quantity of clicks broken down by drift.

::: panel-tabset
## template type=101

```{r}
#| echo: false

template %>%
  filter(type==101) %>%
  count(db)
```

## training

```{r}
#| echo: false
#| 
nbhf_clicks %>%
  count(db) %>%
  mutate(db=str_remove(basename(db), " - Copy.sqlite3"))
```
:::

### Distributions
